maxent model avec xvalidation

### 2019-04-05 src/maxent_avec_xvalidation.Rmd
asc_sum raster looping + terrible loop to make database

@@

```{r final models plus predict}
mth <- "feb"
modelmonth <- cbind(paste0(mth, unique(janprback$year)))
aicetc <- read.csv("../data/env/AICcFinalmod.csv", header = TRUE) #pre-created csv file
unlayer <- read.csv(paste0("../data/env/asc_layerlist", mth, ".csv"), header = TRUE) #if code doesnt work put back to unlayer_yr

@@

      aicetc$Train.AUC <- trainauc
      aicetc$n[i] <- notrainingpoints
      aicetc$coefficients <- nocoeff
      aicetc$sum.of.logs <- sumsamplepred
      aicetc$Test.AUC <- evauc
      aicetc$evcor <- evcor
      aicetc$evpcor <- evpcor
      aicetc$nopresence <- evnp
      aicetc$nobackground <- evna
      aicetc$maxsumsenspec <- evmaxtprtnr
      aicetc$maxsumsenspec <- evmaxkap
      aicetc$no_omission <- evno_omission
      aicetc$equal_sens_spec <- evequal_sens_spec
      aicetc$bic <- bic #not sure this is correct...
      aicetc$beta <- betaval

@@

    unlayer_yr <- subset(unlayer, ascyears == yr)
    unlayer_yr <- unlayer_yr[-c(1), ]
    unlayer_yr[] <- lapply(unlayer_yr, as.character) #from https://stackoverflow.com/questions/2851015/convert-data-frame-columns-from-factors-to-characters
    aea <- raster("../output/env/aea.tif")

    for (td in 1:nrow(unlayer_yr)){
      temp_depth <- raster(unlayer_yr$temp_depth[td])
      crs(temp_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(temp_depth) <- "temp_depth"
      temp_depth <- projectRaster(temp_depth, aea)
      chl_surface <- raster(unlayer_yr$chl_surface[td])
      crs(chl_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(chl_surface) <- "chl_surface"
      chl_surface <- projectRaster(chl_surface, aea)
      o2_depth <- raster(unlayer_yr$o2_depth[td])
      crs(o2_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(o2_depth) <- "o2_depth"
      o2_depth <- projectRaster(o2_depth, aea)
      salinity_depth <- raster(unlayer_yr$salinity_depth[td])
      crs(salinity_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(salinity_depth) <- "salinity_depth"
      salinity_depth <- projectRaster(salinity_depth, aea)
      nao_sample <- raster(unlayer_yr$nao_sample[td])
      crs(nao_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(nao_sample) <- "nao_sample"
      nao_sample <- projectRaster(nao_sample, aea)
      nao_winter <- raster(unlayer_yr$nao_winter[td])
      crs(nao_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
      names(nao_winter) <- "nao_winter"
      nao_winter <- projectRaster(nao_winter, aea)
      stackn <- unlayer_yr$stackname[td]
      stack <- stack(temp_depth, chl_surface, o2_depth, salinity_depth, nao_sample, nao_winter)
      #stackSave(test,unlayer_yr$stackname[td])
      pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mth, "/", stackn, ".asc"), overwrite=TRUE) 
      prplot <- plot(pred, png(paste0("../output/maxent/predictions/", mth, "/", stackn, ".png"))) 
      dev.off() # stops automatic saving of the plot to a png
    }
}
```

@@

```{r avg asc layers}
asc_list <- list.files("../output/maxent/predictions/jan", pattern = '*.asc', full.names = TRUE) #true means the full path is included

>>>>

    #count number of non-zero coefficients, excluding the final four rows
    coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE, stringsAsFactors=FALSE)
    coefile <- head(coefile, -4)
    coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
    nocoeff <- nrow(coefile2)
    bick <- sum(as.numeric(coefile2[, 2])) #dodgy 
    bic <- (bick*log(sum(trainpa == 1)) - 2*11) #dodgy

@@

```{r final models plus predict}
mth <- "jan"
modelmonth <- cbind(paste0(mth, unique(janprback$year)))
aicetc <- read.csv("../data/env/AICcFinalmod.csv", header = TRUE) #pre-created csv file
unlayer <- read.csv(paste0("../data/env/asc_layerlist", mth, ".csv"), header = TRUE) #if code doesnt work put back to unlayer_yr

@@

      aicetc$Train.AUC <- trainauc
      aicetc$n <- notrainingpoints
      aicetc$coefficients <- nocoeff
      aicetc$sum.of.logs <- sumsamplepred
      aicetc$Test.AUC <- evauc
      aicetc$evcor <- evcor
      aicetc$evpcor <- evpcor
      aicetc$nopresence <- evnp
      aicetc$nobackground <- evna
      aicetc$maxsumsenspec <- evmaxtprtnr
      aicetc$maxsumsenspec <- evmaxkap
      aicetc$no_omission <- evno_omission
      aicetc$equal_sens_spec <- evequal_sens_spec
      aicetc$bic <- bic #not sure this is correct...
      aicetc$beta <- betaval

@@

    unlayer_yr <- subset(unlayer, ascyears == yr)
    unlayer_yr <- unlayer_yr[-c(1), ]
    unlayer_yr[] <- lapply(unlayer_yr, as.character) #from https://stackoverflow.com/questions/2851015/convert-data-frame-columns-from-factors-to-characters
    stackdep <- subset(unlayer_yr, select = c(stackname, ascdepths))
    stackdeplst[[year]] <- stackdep
    #write.csv(stackdep, paste0("../output/maxent/predictions/stackdepthlist_", yr, ".csv"), row.names = FALSE)
    aea <- raster("../output/env/aea.tif")

    for (td in 1:nrow(unlayer_yr)){
      temp_depth <- raster(unlayer_yr$temp_depth[td])
      crs(temp_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(temp_depth) <- "temp_depth"
      temp_depth <- projectRaster(temp_depth, aea)
      chl_surface <- raster(unlayer_yr$chl_surface[td])
      crs(chl_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(chl_surface) <- "chl_surface"
      chl_surface <- projectRaster(chl_surface, aea)
      o2_depth <- raster(unlayer_yr$o2_depth[td])
      crs(o2_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(o2_depth) <- "o2_depth"
      o2_depth <- projectRaster(o2_depth, aea)
      salinity_depth <- raster(unlayer_yr$salinity_depth[td])
      crs(salinity_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(salinity_depth) <- "salinity_depth"
      salinity_depth <- projectRaster(salinity_depth, aea)
      nao_sample <- raster(unlayer_yr$nao_sample[td])
      crs(nao_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(nao_sample) <- "nao_sample"
      nao_sample <- projectRaster(nao_sample, aea)
      nao_winter <- raster(unlayer_yr$nao_winter[td])
      crs(nao_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
      names(nao_winter) <- "nao_winter"
      nao_winter <- projectRaster(nao_winter, aea)
      stackn <- unlayer_yr$stackname[td]
      stack <- stack(temp_depth, chl_surface, o2_depth, salinity_depth, nao_sample, nao_winter)
      #stackSave(test,unlayer_yr$stackname[td])
      pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mth, "/", stackn, ".asc"), overwrite=TRUE) 
      prplot <- plot(pred, png(paste0("../output/maxent/predictions/", mth, "/", stackn, ".png"))) 
      dev.off() # stops automatic saving of the plot to a png
    }
}

stdep <- do.call(rbind, stackdeplst)
write.csv(stdep, paste0("../output/maxent/predictions/", mth, "/stack_depth.csv"), row.names = FALSE)
```

@@

```{r avg asc layers}
asc_list <- list.files(paste0("../output/maxent/predictions/", mth), pattern = "*.asc", full.names = TRUE)
depthlayers <- data.frame(cbind(unique(stdep$ascdepths)))
names(depthlayers)[names(depthlayers) == "cbind.unique.stdep.ascdepths.."] <- "depth"
depthlayers$depth <- as.character(depthlayers$depth)

colno <- 1

for (d in 1:nrow(depthlayers)){
  dep <- depthlayers$depth[d]
  for (a in 1:length(asc_list)){
    asc <- asc_list[a]
    if (grepl(dep, asc)){
      depthlayers[d , paste0("a", colno)] <- asc
        colno <- colno+1
    }
  }
}
write.csv(depthlayers, (paste0("../output/maxent/predictions/", mth, "/asclayerbydepth.csv")), row.names = TRUE)

for (x in 1:nrow(depthlayers)){
  asclay <- depthlayers[1, ]
  asclay <- asclay[!is.na(asclay)]
  asclay <- as.data.frame(rbind(asclay))
  asclay$V1 <- as.character(asclay$V1)
  asclay$V2 <- as.character(asclay$V2)
  asclay$V3 <- as.character(asclay$V3)
  asclay$V4 <- as.character(asclay$V4)
  asclay$V5 <- as.character(asclay$V5)
  asclaydep <- asclay$V1

  asc2003 <- raster(asclay$V2)
  asc2004 <- raster(asclay$V3)
  asc2005 <- raster(asclay$V4)
  asc2006 <- raster(asclay$V5)

  stkasc <- stack(asc2003, asc2004, asc2005, asc2006)
  avg <- calc(stkasc, fun = mean)

  writeRaster(avg, (paste0("../output/maxent/predictions/", mth, "/avg_", asclaydep, ".asc")), overwrite=TRUE)
  avgplot <- plot(avg, png(paste0("../output/maxent/predictions/", mth, "/avg_", asclaydep, ".png"))) 
  dev.off()
}
